---
title: "Inclusive Teaching Text Analysis - Part 2"
author: "Yuhao Zhao"
date: "2022-12-14"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Setup/Data Cleaning

Here, I've imported the excel data set and necessary packages for analysis. I also did some data cleaning, created a vector of DEI related words and added some variables to the original excel data set.

```{r keyboard shortcut}

#cmd + shift + c to comment out code 
#cmd + shift + M to print %>% pipe operator
#cmd + return to run code 

```

```{r Selecting packages}
# Store string containing all required packages
my_packages <- c('varhandle', 'skimr', 'tidyverse', 'tidytext', 'stopwords', "wordcloud", "reshape2", "ggraph", "kableExtra",'readr', 'dplyr', "igraph","SnowballC")
```

Figure out which of these packages is already installed

```{r Check package status}
# Store all installed packages
ya_installed <- library()$results[,1]
# Check whether required packages are already installed and grab only those that still need installation
need_install<-my_packages[!(my_packages %in% ya_installed)]
#install required packages
lapply(need_install, install.packages, character.only = TRUE)
```

Now, load only unloaded packages

```{r Loading packages and data}
# Store all installed packages
ya_loaded <- (.packages())
# Check whether required packages are already installed and grab only those that still need installation
need_load<-my_packages[!(my_packages %in% ya_loaded)]
# Load required packages
lapply(need_load, require, character.only = TRUE)
```




## Network Plot of Word Relationship over Time {.tabset}

To get a deeper understanding of how inclusive teaching is viewed, we will be creating a network plot to look at the relationship between words/phrases in the inclusive teaching section. The generated `igraph` graph is called `rios_phrase_network`. It has 41 words and 36 connections among them. Similar to what some of our graphs have portrayed above, the words "inclusive", "students", and "diverse" are connected to many other words.

### Datasets perparation
```{r}
updated_DEI<- read_csv("/Users/zhaoyuhao/Desktop/College/Research/RIOS---CourseSource-Research/Data Cleaning/Data for Export/JEDIkeywords.csv")
updated_2DEI <- as.data.frame(read_csv("/Users/zhaoyuhao/Desktop/College/Research/RIOS---CourseSource-Research/Data Cleaning/Data for Import/cleanedIT2words.csv", show_col_types = FALSE)[1:3])
rios_data_token2it <- read_csv("/Users/zhaoyuhao/Desktop/College/Research/RIOS---CourseSource-Research/Data Cleaning/Data for Export/rios_data_tokenized2.csv")
# seperate the two word pairs coloumn into two seperate ones and subsetting by verified lists
rios_data_token2 <- rios_data_token2it %>% filter(it_tokens_2w %in% updated_2DEI$it_tokens_2w) %>% separate(it_tokens_2w, c("word1", "word2"), sep = " ") 

```


### 2014

```{r 2014}

#extracting the edgelist(two-word pairs) and create a igraph object from it
rios_2w_net_2014.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2014) %>%
  count(word1, word2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2014.g)[V(rios_2w_net_2014.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2014.g)[!V(rios_2w_net_2014.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2014_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2014) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2014_stemmed.g)[V(rios_2w_net_2014_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2014_stemmed.g)[!V(rios_2w_net_2014_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"
```


```{r 2014}
#Visualization
set.seed(20181005)

#original net
png(file="2w2014.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2014.g, layout=layout_with_kk, edge.width =E(rios_2w_net_2014.g)$n, vertex.size=log(degree(rios_2w_net_2014.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2014")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

#stemmed net
png(file="2w2014_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2014_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2014_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2014_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2014")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()
```

### 2015

```{r 2015}

#create a igraph graph object
rios_2w_net_2015.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2015) %>%
  count(word1, word2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2015.g)[V(rios_2w_net_2015.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2015.g)[!V(rios_2w_net_2015.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2015_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2015) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2015_stemmed.g)[V(rios_2w_net_2015_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2015_stemmed.g)[!V(rios_2w_net_2015_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"
```


```{r 2015}
#Visualization
set.seed(20181005)

#original net
png(file="2w2015.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2015.g, layout=layout_with_kk, edge.width =E(rios_2w_net_2015.g)$n, vertex.size=log(degree(rios_2w_net_2015.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2015")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

#stemmed version
png(file="2w2015_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2015_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2015_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2015_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2015")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()
```


### 2016


```{r 2016}
# rios_data_token2 <- rios_data_token2it %>%
#   separate(it_tokens_2w, c("word1", "word2"), sep = " ")

rios_2w_net_2016.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2016) %>%
  count(word1, word2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2016.g)[V(rios_2w_net_2016.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2016.g)[!V(rios_2w_net_2016.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2016_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2016) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2016_stemmed.g)[V(rios_2w_net_2016_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2016_stemmed.g)[!V(rios_2w_net_2016_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"
```


```{r 2016}
set.seed(20181005)


#original net
png(file="2w2016.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2016.g, layout=layout_with_kk, edge.width =E(rios_2w_net_2016.g)$n, vertex.size=log(degree(rios_2w_net_2016.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2016")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

#stemmed net
png(file="2w2016_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2016_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2016_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2016_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2016")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()
```

### 2017

```{r 2017}

rios_2w_net_2017.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2017) %>%
  count(word1, word2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2017.g)[V(rios_2w_net_2017.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2017.g)[!V(rios_2w_net_2017.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2017_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2017) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2017_stemmed.g)[V(rios_2w_net_2017_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2017_stemmed.g)[!V(rios_2w_net_2017_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"
```


```{r 2017}
set.seed(20181005)

png(file="2w2017.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2017.g, layout=layout_with_fr, edge.width =E(rios_2w_net_2017.g)$n, vertex.size=log(degree(rios_2w_net_2017.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2017")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

png(file="2w2017_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2017_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2017_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2017_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2017")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()
```

### 2018

```{r 2018}

rios_2w_net_2018.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2018) %>%
  count(word1, word2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2018.g)[V(rios_2w_net_2018.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2018.g)[!V(rios_2w_net_2018.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2018_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2018) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2018_stemmed.g)[V(rios_2w_net_2018_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2018_stemmed.g)[!V(rios_2w_net_2018_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"
```


```{r 2018}
set.seed(20181005)

png(file="2w2018.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2018.g, layout=layout_with_kk, edge.width =E(rios_2w_net_2018.g)$n, vertex.size=log(degree(rios_2w_net_2018.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2018")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

png(file="2w2018_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2018_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2018_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2018_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2018")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()
```

### 2019

```{r 2019}

rios_2w_net_2019.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2019) %>%
  count(word1, word2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2019.g)[V(rios_2w_net_2019.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2019.g)[!V(rios_2w_net_2019.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2019_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2019) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2019_stemmed.g)[V(rios_2w_net_2019_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2019_stemmed.g)[!V(rios_2w_net_2019_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

```


```{r 2019}
set.seed(20181005)

#original net
png(file="2w2019.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2019.g, layout=layout_with_kk, edge.width =E(rios_2w_net_2019.g)$n, vertex.size=log(degree(rios_2w_net_2019.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2019")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

#stemmed version
png(file="2w2019_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)
plot(rios_2w_net_2019_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2019_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2019_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2019")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

```


### 2020

```{r 2020}
rios_2w_net_2020.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2020) %>%
  count(word1, word2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2020.g)[V(rios_2w_net_2020.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2020.g)[!V(rios_2w_net_2020.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2020_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2020) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2020_stemmed.g)[V(rios_2w_net_2020_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2020_stemmed.g)[!V(rios_2w_net_2020_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"
```


```{r 2020}
set.seed(20181005)

#original net
png(file="2w2020.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2020.g, layout=layout_with_kk, edge.width =E(rios_2w_net_2020.g)$n*2, vertex.size=log(degree(rios_2w_net_2020.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2020")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

png(file="2w2020_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2020_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2020_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2020_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2020")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()
```

### 2021

```{r 2021}
rios_2w_net_2021.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2021) %>%
  count(word1, word2, sort = TRUE) %>%  
  graph_from_data_frame()

V(rios_2w_net_2021.g)[V(rios_2w_net_2021.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2021.g)[!V(rios_2w_net_2021.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2021_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2021) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2021_stemmed.g)[V(rios_2w_net_2021_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2021_stemmed.g)[!V(rios_2w_net_2021_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"
```


```{r 2021}
set.seed(20181005)


#original net
png(file="2w2021.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2021.g, layout=layout_with_kk, edge.width =E(rios_2w_net_2021.g)$n, vertex.size=log(degree(rios_2w_net_2021.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2021")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

#stemmed version
png(file="2w2021_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2021_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2021_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2021_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2021")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()
```

### 2022

```{r 2022}
rios_2w_net_2022.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2022) %>%
  count(word1, word2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2022.g)[V(rios_2w_net_2022.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2022.g)[!V(rios_2w_net_2022.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"

#stemmed version
rios_2w_net_2022_stemmed.g <- rios_data_token2 %>% 
  filter(dei_related == TRUE & Year == 2022) %>%
  mutate(stem1 = wordStem(`word1`, language = "en")) %>%
  mutate(stem2 = wordStem(`word2`, language = "en")) %>%
  count(stem1, stem2, sort = TRUE) %>% 
  graph_from_data_frame()

V(rios_2w_net_2022_stemmed.g)[V(rios_2w_net_2022_stemmed.g)$name %in% updated_DEI$allwords]$color<-"#1f78b4"
V(rios_2w_net_2022_stemmed.g)[!V(rios_2w_net_2022_stemmed.g)$name %in% updated_DEI$allwords]$color<- "#a6cee3"
```


```{r 2022}
set.seed(20181005)


png(file="2w2022.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2022.g, layout=layout_with_kk, edge.width =E(rios_2w_net_2022.g)$n, vertex.size=log(degree(rios_2w_net_2022.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) Word Relationship in 2022")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()

png(file="2w2022_stemmed.png",res=1000, width=10000, height=10000, pointsize = 9)

plot(rios_2w_net_2022_stemmed.g,layout=layout_with_kk, edge.width =E(rios_2w_net_2022_stemmed.g)$n, vertex.size=log(degree(rios_2w_net_2022_stemmed.g)),vertex.label.degree=-pi/2,
vertex.label.dist=0.2,vertex.label.cex=0.3, edge.arrow.size=0.3, edge.arrow.width=0.2, main="Network of (DEI Related) stemmed Word Relationship in 2022")

legend(0.95, 1, legend=c("2DEI","1DEI")  , col = c("#a6cee3", "#1f78b4") , bty = "n", pch=c(20, 20) , lty =c(0, 0),  pt.cex = 2, cex = 1, text.col=c("#a6cee3", "#1f78b4") ,xjust = 0, yjust = 1, x.intersp = 1, y.intersp = 1,adj = c(0, 0.5), horiz = FALSE)
dev.off()
```


### Statistical measure summary

#### create a graph list
```{r}

graph_list<-list(rios_2w_net_2014.g, rios_2w_net_2014_stemmed.g, rios_2w_net_2015.g, rios_2w_net_2015_stemmed.g, rios_2w_net_2016.g, rios_2w_net_2016_stemmed.g, rios_2w_net_2017.g, rios_2w_net_2017_stemmed.g, rios_2w_net_2018.g, rios_2w_net_2018_stemmed.g, rios_2w_net_2019.g, rios_2w_net_2019_stemmed.g, rios_2w_net_2020.g, rios_2w_net_2020_stemmed.g, rios_2w_net_2021.g, rios_2w_net_2021_stemmed.g, rios_2w_net_2022.g, rios_2w_net_2022_stemmed.g)
```

```{r}
#network size
sizes<-sapply(graph_list, gorder)
   
#edge density

conns<-sapply(graph_list, edge_density)

#Diameter

diameters<-sapply(graph_list, diameter)

#clustering coefficient

ccoeffs<-sapply(graph_list, transitivity)

#components
n_comps<-sapply(graph_list, count_components)

#max component size
max_comp_size<-sapply(sapply(graph_list, components)[2, ], max)

#Summary stats for networks across years
summ_df_new<-rbind(sizes, conns, diameters, ccoeffs, n_comps, max_comp_size)

colnames(summ_df_new)<-c("2014", "2014st", "2015", "2015st", "2016", "2016st", "2017", "2017st", "2018", "2018st", "2019", "2019st", "2020", "2020st", "2021", "2021st", "2022", "2022st")
summ_df_new<-t(summ_df_new)

View(summ_df_new)
```
